Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-08-01 draft.tex
continued main section

\subsubsection{Assessing the Information Gain Using the PCA}
%%
To determine the information gain per image-derived variables, we calculated a PCA and compared the loadings per dimension.
%%
The first three principal components explain 75\% of variance in the image-derived variables.
%%
The first principal component explains 47\% of the variance and weights primarily \emph{mean curvature}, \emph{curvature sagittal}, \emph{curvature angle} and \emph{curvature angle sagittal}.
%%
The second component, adding 16\% of variance, weigths  \emph{curvature coronal} and \emph{curvature angle coronal}.
%%
The third component explains 12\% of variance and weights \emph{torsion} and \emph{curvature transverse}.
%%
This supports our prior conclusion about the low information gain of the transverse planes.
%%
\emph{Torsion} also adds little variance to the information space.
%% ------------------------------------------
\subsubsection{Heterogenous Correlations}
%%
We then expanded our focus on correlations of image-derived parameters with all other non-image variables.
%%
We applied a heterogenous correlation technique to derive correlations between all variables is the data set.
%%
The method uses the following correlation metrics for the different type combination:
%%
\begin{itemize}
%%
\item \emph{pearson product-moment} for two continuous variables, 
\item \emph{polyserial} correlation for one continuous and one categorical variable, and
\item \emph{polychoric} correlations for two categorical variables.
\end{itemize}
%%
All correlation values are scaled between 0 - no correlation and 1 - identical.
%%
Some variables are too sparse for calculating correlations, for example \com{some variables from calculation here}.
%%
We display the resulting \emph{contingency matrix} using a heat map, mapping correlation values to color brightness with white for 0 and dark blue for 1 \cite{Klemm2014VIS}.
%%
We calculated the contingency matrix for all size groups and looked for correlations between image- and non-image derived variables.
%%
The resulting contingency matrix shows no strong correlation with any of the parameter.
%%
Only weak correlations could be found for \emph{Mean Curvature} with \emph{gender} (0.42), \emph{body size} (0.39) and \emph{number of born children} (0.29).
%%
One surprising result was the small correlation with \emph{torsion}, which correlated with almost no variables (p values between 0 and 0.05) and \emph{parkinson} (0.24).
%%
This observation brought us to the conclusion to incorporate more sophisticated data mining techniques to assess the influence of the image-derived parameters.

@@

\subsection{Interactive Display of Decision Trees}
\com{Alle Variablen einzeln durchgehen ist nervig, man braucht ein interaktives System hierfür!}
\com{Describe Shiny App!}

\subsection{Das hier unterbringen}
\begin{itemize}
\item \com{Vorgehen mit Scatterplot erklären}
\item \com{Beschreiben der Ergebnisse}
\item \com{Beschreiben des interaktiven Systems zum Anzeigen der Decision Trees}
\end{itemize}

>>>>

\subsubsection{Assessing the Information Gain Using the PCA}
%%
To determine the information gain per image-derived variables, we calculated a PCA and compared the loadings per dimension.
%%
The first three principal components explain 75\% of variance in the image-derived variables.
%%
The first principal component explains 47\% of the variance and weights primarily \emph{mean curvature}, \emph{curvature sagittal}, \emph{curvature angle} and \emph{curvature angle sagittal}.
%%
The second component, adding 16\% of variance, weigths  \emph{curvature coronal} and \emph{curvature angle coronal}.
%%
The third component explains 12\% of variance and weights \emph{torsion} and \emph{curvature transverse}.
%%
This supports our prior conclusion about the low information gain of the transverse planes.
%%
\emph{Torsion} also adds little variance to the information space.
%% ------------------------------------------
\subsubsection{Heterogenous Correlations}
%%
We then expanded our focus on correlations of image-derived parameters with all other non-image variables.
%%
We applied a heterogenous correlation technique to derive correlations between all variables is the data set.
%%
The method uses the following correlation metrics for the different type combination:
%%
\begin{itemize}
%%
\item \emph{pearson product-moment} for two continuous variables, 
\item \emph{polyserial} correlation for one continuous and one categorical variable, and
\item \emph{polychoric} correlations for two categorical variables.
\end{itemize}
%%
All correlation values are scaled between 0 - no correlation and 1 - identical.
%%
Some variables are too sparse for calculating correlations, for example \com{TODO: some variables from calculation here; TODO: Image?, eventuell interactiven Tile-Plot}.
%%
We display the resulting \emph{contingency matrix} using a heat map, mapping correlation values to color brightness with white for 0 and dark blue for 1 \cite{Klemm2014VIS}.
%%
We calculated the contingency matrix for all size groups and looked for correlations between image- and non-image derived variables.
%%
The resulting contingency matrix shows no strong correlation with any of the parameter.
%%
Only weak correlations could be found for \emph{Mean Curvature} with \emph{gender} (0.42), \emph{body size} (0.39) and \emph{number of born children} (0.29).
%%
One surprising result was the small correlation with \emph{torsion}, which correlated with almost no variables (p values between 0 and 0.05) and \emph{parkinson} (0.24).
%%
This observation brought us to the conclusion to incorporate more sophisticated data mining techniques to assess the influence of the image-derived parameters.

@@

\subsection{Interactive Display of Decision Trees}
We have to create a decision tree for every non-image variable to see which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields a corresponding number of trees.
%%
Dividing the subjects into four groups using body-size, the number increases to 402 trees.
%%
To make the results cognitive feasible, we have to abstract the results of the classification.
%%------------
\subsubsection{Visualization of Classification Results}
%%
We follow the visual information-seeking mantra and want to provide an overview first, then details-on-demand \ref{shneiderman1996}.
%%
The optimal classification uses a few rules to precisely characterize the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error rate}.
%%
The error rate has to be judged depending on the number of classes of the target variable.
%%
An error rate of 50\% for a target variable with two manifestations makes the classifier equal to throwing a coin.
%%
The same error rate is acceptable a target variable with more than two manifestations.
%%
Picking a value for a variable of four manifestations with an error rate of 50\% is better than the random selection (error rate of 75\%).
%%
The following scales the error rate so that 0 denotes perfect classification and 1 is equal to random selection.
%%
\begin{equation}
error_{normalized} = \frac{error}{100 - \frac{1}{class number}}
\end{equation}
%%
This allows for comparability of error rates between variables.
%%
Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatterplot.
%%
By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%
\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We however still want to be able to look display \emph{details-on-demand} \ref{shneiderman1996} and examinate a decision tree in detail.
%%
This is realized by clicking on a entry on the visualization, which then displays the corresponding decision tree in detail (Fig.~\ref{fig:decision-tree} b).
%%
This allows for sequentially analyzing the interactions.
%%------------
\subsubsection{Implementation}
All analysis are carried out using \texttt{R}, a popular programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.
%%
As opposed to the standard \texttt{R} plots, \texttt{ggvis} allows to adjust visualization parameters using UI controls, such as sliders.
%%
We implemented our application as web based prototype using \texttt{R Shiny}\footnote{R Shiny footnote}.
%%
The package combines the power of \texttt{R} with advantages of web-based applications.
%%
It allows us to quickly exchange results with our collaborating epidemiological partners.
%%
They can try out the technique without installing anything.
%%
Exchanging software becomes as easy as exchanging a hyperlink.
%%---------
\subsection{Results}

\subsection{Themen hier unterbringen}
